---
title: "ECD Connect Data Report"
author: "Kim Tichmann"
format: html
---

# October 2025 Report
Data report as of 31 October 2025

## ECD Connect in the ECD Data Landscape

South Africa’s ECD sector is moving towards a more integrated data system, with efforts underway to strengthen the DBE’s EMIS and align partner data contributions.

ECD Connect plays a key role in this ecosystem - not only supporting practitioners and CHWs on the ground, but also generating rich data about early learning and health service delivery.

This report provides an initial snapshot of some of the data emerging from ECD Connect and explores how, as the datasets grow, they could contribute to broader insight and learning across the sector.
<br>

---

<br>

## 1. Reach and Scale
ECD Connect has grown rapidly across both education and health since launch. The figures below show our overall footprint to date, reflecting adoption across the open-access app, partner versions, and Grow Great’s CHW Connect.

The totals represent cumulative numbers since each platform’s laun


::: {.stat-block}

::: {.stat-card}
<p class="stat-number">4</p>
<p class="stat-label">partner organisations</p>
:::

::: {.stat-card}
<p class="stat-number">2,160</p>
<p class="stat-label">unique user sign-ups</p>
:::

::: {.stat-card}
<p class="stat-number">12,703</p>
<p class="stat-label">total number of children added across all platforms</p>
:::

:::

```{r beneficiaries-map, echo=FALSE}
#| message: false
#| warning: false
#| paged-print: true

library(tidyverse)
library(sf)
library(dplyr)
library(ggplot2)
library(conflicted)
library(stringr)
library(patchwork)
library(scales)
library(viridis)
library(lubridate)
library(readr)
library(leaflet)
library(plotly)
library(gt)
library(tidyr)
library(knitr)
library(kableExtra)
library(gtExtras)

conflict_prefer("filter","dplyr")

load("~/Desktop/data/sub_places.rda")  
load("~/Desktop/data/postalcodes_geo.rda")

df_children <- read.csv("~/Desktop/data/codes-ed.csv", stringsAsFactors = FALSE)
df_clients  <- read.csv("~/Desktop/data/GG-location.csv", stringsAsFactors = FALSE)

districts <- suppressMessages(suppressWarnings(st_read("~/Desktop/data/DistrictMunicipalities2018_Final.shp", quiet = TRUE)))

df_clients2 <- df_clients %>%
  mutate(
    District_clean = trimws(toupper(District))
  )

districts2 <- districts %>%
  mutate(
    DISTRICT_NAME_clean = trimws(toupper(DISTRICT_N))
  )

# Which CSV district names are in the map?
matched <- df_clients2 %>%
  filter(District_clean %in% districts2$DISTRICT_NAME_clean)
invisible(nrow(matched))

# ensure the district polygon column exists
if (!"DISTRICT_NAME_clean" %in% names(districts2)) {
  stop("Expected DISTRICT_NAME_clean in 'districts2' but not found.")
}

# Join the health data to the district polygons
districts_health <- districts2 %>%
  left_join(df_clients2, by = c("DISTRICT_NAME_clean" = "District_clean"))

df_children2 <- df_children %>%
  mutate(PostalCode = as.character(PostalCode),
         TotalChildren = as.numeric(TotalChildren))

postalcodes_geo2 <- postalcodes_geo %>%
  mutate(PostalCode = as.character(code))

# join lat/lon + totals
pcs <- postalcodes_geo2 %>%
  left_join(df_children2, by = "PostalCode") %>%
  filter(!is.na(TotalChildren))

# convert to spatial points
pcs_sf <- st_as_sf(pcs, coords = c("lon", "lat"), crs = 4326, remove = FALSE)

# align CRS + spatial join to districts
pcs_sf <- st_transform(pcs_sf, st_crs(districts2))
pcs_with_district <- st_join(pcs_sf, districts2 %>% select(DISTRICT_NAME_clean), join = st_within)

# aggregate postal codes up to district
districts_ed <- pcs_with_district %>%
  st_set_geometry(NULL) %>%
  group_by(DISTRICT_NAME_clean) %>%
  summarise(TotalChildren = sum(TotalChildren, na.rm = TRUE)) %>%
  right_join(districts2, by = "DISTRICT_NAME_clean") %>%
  st_as_sf()

# ---- COMBINE BOTH LAYERS ----
health_for_map <- districts_health %>%
  select(DISTRICT_NAME_clean, geometry, Value = TotalClients) %>%
  mutate(Source = "Health")

ed_for_map <- districts_ed %>%
  select(DISTRICT_NAME_clean, geometry, Value = TotalChildren) %>%
  mutate(Source = "Education")

combined_map <- bind_rows(health_for_map, ed_for_map) %>%
  group_by(DISTRICT_NAME_clean) %>%
  summarise(
    TotalBeneficiaries = sum(Value, na.rm = TRUE),
    geometry = sf::st_union(geometry)
  ) %>%
  ungroup()

# --- Merge with district polygons to ensure full geometry coverage ---
combined_map <- districts2 %>%
  left_join(
    st_drop_geometry(combined_map),
    by = "DISTRICT_NAME_clean"
  )


# --- Plot ---

combined_map <- combined_map %>%
  mutate(
    BeneficiaryBand = case_when(
      is.na(TotalBeneficiaries) ~ "0",
      TotalBeneficiaries == 0 ~ "0",
      TotalBeneficiaries <= 500 ~ "1–500",
      TotalBeneficiaries <= 1000 ~ "501–1000",
      TotalBeneficiaries <= 1500 ~ "1001–1500",
      TotalBeneficiaries <= 2000 ~ "1501–2000",
      TotalBeneficiaries <= 2500 ~ "2001–2500",
      TotalBeneficiaries <= 3000 ~ "2501–3000",
      TotalBeneficiaries <= 3500 ~ "3001–3500",
      TotalBeneficiaries <= 4000 ~ "3501–4000",
      TotalBeneficiaries > 4000 ~ "4000+"
    ),
    BeneficiaryBand = factor(
      BeneficiaryBand,
      levels = c("0", "1–500", "501–1000", "1001–1500", "1501–2000",
                 "2001–2500", "2501–3000", "3001–3500", "3501–4000", "4000+")
    )
  )


brand_palette <- c(
  "0" = "#EFEFEF",        # light grey
  "1–500" = "#D8EBF6",    # very light blue
  "501–1000" = "#CBE9F4",
  "1001–1500" = "#A4DBEF",
  "1501–2000" = "#8EDCEF",
  "2001–2500" = "#5CCAE2",
  "2501–3000" = "#1DBADF",
  "3001–3500" = "#1497BF",
  "3501–4000" = "#52607B",
  "4000+" = "#27385A"     # darkest brand blue
)

# Create dummy rows for each BeneficiaryBand
dummy_levels <- data.frame(
  BeneficiaryBand = factor(
    names(brand_palette),
    levels = names(brand_palette)
  )
)

# Create dummy geometries with the same CRS as your combined_map
dummy_levels$geometry <- sf::st_sfc(
  replicate(
    length(brand_palette),
    sf::st_point(c(0, 0)),
    simplify = FALSE
  ),
  crs = sf::st_crs(combined_map)   # ensure same CRS
)

# Convert dummy_levels into an sf object
dummy_levels <- sf::st_as_sf(dummy_levels)

# Combine with your real map data
combined_map2 <- dplyr::bind_rows(
  combined_map %>% select(BeneficiaryBand, geometry),
  dummy_levels
)


ggplot(combined_map2) +
    geom_sf(aes(fill = BeneficiaryBand), colour = "#fff", size = 0.15) +
  scale_fill_manual(
    values = brand_palette,
    name = "Total beneficiaries",
    drop = FALSE,
    na.value = brand_palette["0"],
    guide=guide_legend(
      override.aes = list(fill = unname(brand_palette))
    )
  ) +
  coord_sf(xlim = st_bbox(combined_map)[c("xmin", "xmax")],
         ylim = st_bbox(combined_map)[c("ymin", "ymax")],
         expand = FALSE) +
  labs(
    title = "Total Beneficiaries by District",
    subtitle = "Combined Health and Education data",
    caption = "‘Beneficiaries’ refers to children (education and health) \n and pregnant clients (health) registered in the app."
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

```

Together, these numbers represent the growing network of practitioners, CHWs, and children engaging with the platform.

---

<br>

## 2. Versions and Usage Patterns

ECD Connect operates through several tailored versions, each designed for specific user groups. The sections below summarise usage and engagement across education and health.

### Education
#### ECD Connect App
_An open-access version of the app - open to any principal/practitioner working in ECD in South Africa. Since launch in April 2025 to October 2025_

::: {.stat-block}

::: {.stat-card .app-background}
<p class="stat-number">1,605</p>
<p class="stat-label">unique sign-ups</p>
:::

::: {.stat-card .app-background}
<p class="stat-number">537</p>
<p class="stat-label">preschools added</p>
:::

::: {.stat-card .app-background}
<p class="stat-number">1,637</p>
<p class="stat-label">children registered</p>
:::

:::

##### User Retention

- **17%** of users who joined the app over 30 days ago have logged in again after 30 days. This is higher than the standard [global app retention rate of 7% by day 30](https://www.adjust.com/resources/guides/user-retention/).
- **8%** of practitioners who signed up over 30 days ago have logged in within the past month.


#### ECD Connect Partner
_A customisable version for organisations who want to use a organisation-branded version of the app within their networks. We currently have versions for: Khululeka, Ntataise, and True North. Since pilot launch in October 2024 to October 2025_

::: {.stat-block}

::: {.stat-card .partner-background}
<p class="stat-number">226</p>
<p class="stat-label">unique sign-ups</p>
:::

::: {.stat-card .partner-background}
<p class="stat-number">61</p>
<p class="stat-label">preschools added</p>
:::

::: {.stat-card .partner-background}
<p class="stat-number">1,020</p>
<p class="stat-label">children registered</p>
:::

:::

##### User Retention

- **26%** of users who joined the app over 30 days ago have logged in again after 30 days.
- **12%** of practitioners who signed up over 30 days ago have logged in within the past month.

### Health

#### CHW Connect
_Since launch in April 2024 to October 2025_

::: {.stat-block}

::: {.stat-card .gg-background}
<p class="stat-number">342</p>
<p class="stat-label">CHWs registered</p>
:::

::: {.stat-card .gg-background}
<p class="stat-number">60 202</p>
<p class="stat-label">child visits completed</p>
:::

::: {.stat-card .gg-background}
<p class="stat-number">6 501</p>
<p class="stat-label">pregnant client visits completed</p>
:::

:::

##### User Retention

- **79%** of CHWs who joined the app over 30 days ago have logged in again after 30 days.
- **35%** of CHWs who signed up over 30 days ago have logged in within the past month.

<br>

---

<br>

## 3. Emerging Insights
_As use of the education app expands, richer patterns are emerging in the data being captured — from child demographics and attendance to preschool business trends. The visuals below provide early insight into how the platform reflects real-world programme participation and operations._

### Education

#### 1. Child Demographics

##### Geographic Spread of Registered Children

```{r ed-children-map, echo=FALSE}
#| message: false
#| warning: false
#| paged-print: true

# --- Load shapefile ---
districts <- st_read("~/Desktop/data/DistrictMunicipalities2018_Final.shp", quiet = TRUE) %>%
  st_transform(4326)

# --- Load CSV ---
codes <- read_csv("~/Desktop/data/codes-ed.csv")

# --- Merge with postal code coordinates ---
codes_geo <- codes |>
  left_join(postalcodes_geo, by = c("PostalCode" = "code")) |>
  drop_na(lat, lon)

# --- Convert to spatial points ---
codes_sf <- st_as_sf(codes_geo, coords = c("lon", "lat"), crs = 4326)

# --- Create ggplot (static) ---
p <- ggplot() +
  geom_sf(data = districts, fill = "grey95", colour = "grey70", linewidth = 0.3) +
  geom_sf(
    data = codes_sf,
    aes(size = TotalChildren,
        text = paste0("Postal code: ", PostalCode,
                      "<br>Total children: ", TotalChildren)),
    colour = "#0072B2", alpha = 0.6
  ) +
  scale_size_continuous(range = c(1, 10)) +
  coord_sf(datum = NA) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(title = "Total Children by Postal Code") +
  guides(size = "none")

# --- Make interactive map ---
leaflet() |>
  addProviderTiles("CartoDB.Positron") |>
  addPolygons(
    data = districts,
    color = "#666", weight = 1, fillOpacity = 0.1
  ) |>
  addCircleMarkers(
    data = codes_sf,
    radius = ~scales::rescale(TotalChildren, to = c(2, 12)),
    color = "#1DBADF", fillOpacity = 0.8, stroke = FALSE,
    popup = ~paste0("Total children: ", TotalChildren)
  ) |>
  setView(lng = 25, lat = -29, zoom = 5) |>
  setMaxBounds(
    lng1 = 16, lat1 = -35,   # Southwest corner (approx)
    lng2 = 33, lat2 = -22    # Northeast corner (approx)
  )

```

<br>



<div class="chart-row">

<div class="chart-box">

##### Child Gender Distribution

``` {r ed-children-gender, echo=FALSE }
#| warning: false
#| message: false

conflicted::conflicts_prefer(plotly::layout)

# Load data
gender_data <- read_csv("~/Desktop/data/child-gender.csv")

# Prepare data
gender_data <- gender_data %>%
  arrange(desc(count)) %>%
  mutate(
    gender = factor(gender, levels = gender),
    Percentage = round(count / sum(count) * 100, 1)
  )

# Create interactive pie
plot_ly(
  data = gender_data,
  labels = ~gender,
  values = ~count,
  type = "pie",
  sort = FALSE,
  direction = "clockwise",
  textinfo = "label+percent",
  textposition = "auto",
  hoverinfo = "text",
  text = ~paste(gender, ": ", Percentage, "%"),
  marker = list(
    colors = c("#83BB26", "#1DBADF", "#FFD525"),
    line = list(color = "white", width = 2)
  )
) %>%
  layout(
  showlegend = FALSE,
  uniformtext = list(minsize = 12, mode = "hide"),
  startangle = 90,
  margin = list(l = 0, r = 0, t = 10, b = 0),
  automargin = TRUE,
  autosize = TRUE,
  height = NULL,
  width = NULL,
  piecolorway = c("#83BB26", "#1DBADF", "#FFD525"),
  annotations = list(),
  # 👇 this is what enables the leader (connector) lines
  showlegend = FALSE,
  # Important connector line setting:
  newshape = list(line = list(color = "grey", width = 1))
) %>%
  config(responsive = TRUE)


```
</div>

<div class="chart-box">
##### Child Age Distribution

``` {r ed-children-ages, echo=FALSE, fig.height = 8 }
#| warning: false
#| message: false

# Read the CSV

age_data <- read_csv("~/Desktop/data/age-groups-ed.csv")

# Create the bar chart
ggplot(age_data, aes(x = `Age Group`, y = `Number of Children`)) +
geom_col(fill = "#1DBADF") +
labs(x = "Age Group",
y = "Number of children") +
theme_minimal(base_size = 13) +
theme(
plot.title = element_text(face = "bold", hjust = 0.5),
axis.text.x = element_text(angle = 30, hjust = 1)
)
```

<br>


</div></div>

##### Family Grant Types

```{r grants-ed, echo=FALSE}
#| warning: false
#| message: false

# Load data
grants <- read_csv("~/Desktop/data/grants-ed.csv")

# Clean up labels
grants <- grants %>%
  mutate(type = recode(type,
                       "None selected" = "No response"))

# Plot
ggplot(grants, aes(x = fct_reorder(type, number_of_children), 
                   y = number_of_children)) +
  geom_col(fill = "#1DBADF", show.legend = FALSE) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Number of children"
  ) +
  theme_minimal(base_size = 13)
```

<br>

#### 2. Access & Participation

##### Attendance registers saved
_As we roll out to more users, we are seeing an increase in the number of attendance registers saved per month._

```{r app-attendance-bar, echo=FALSE}
#| message: false
#| warning: false
#| paged-print: true
#| out.width: 100%

df_appattendance <- read.csv("~/Desktop/data/app-attendance.csv", stringsAsFactors = FALSE)

df_appattendance <- df_appattendance %>%
  mutate(
    # Parse your "YYYY-MM" into a proper Date
    MonthDate = ym(Month)  # lubridate::ym() interprets "2025-04" as 2025-Apr-01
  )

ggplot(df_appattendance, aes(x = MonthDate, y = Registers.Saved)) +
  geom_col(fill = "#FF2180") +
  scale_x_date(
  date_labels = "%b %Y",   # 👈 shows "Apr 2025" instead of "April 2025"
  date_breaks = "1 month"
) +
  labs(
    x = "Month",
    y = "Number of registers saved"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

<br>

##### Average Child Attendance by Month

```{r child-attendance, echo=FALSE}
#| warning: false
#| message: false

# --- Load and prepare data ---
attendance <- read_csv("~/Desktop/data/child-attendance.csv") %>%
  arrange(month_start)

# --- Plot ---
ggplot(attendance, aes(x = reorder(Month, month_start),
                       y = `Percentage child attendance`)) +
  geom_col(fill = "#1DBADF") +                           # ECD Connect teal
  geom_text(aes(label = paste0(round(`Percentage child attendance`, 0), "%")),
            vjust = -0.5, size = 3.2, colour = "#004C6D") +  # navy text
  scale_y_continuous(labels = label_percent(scale = 1)) +
  labs(
    x = NULL,
    y = "Child attendance (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", colour = "#004C6D"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )
```

<br>

#### 3. Business & Finance

##### Spread of preschool balances (1 August to 31 October 2025)
_Balance is income minus expenses_

```{r balances-table, echo=FALSE}
#| warning: false
#| message: false

options(readr.show_col_types = FALSE)

invisible (balances <- suppressMessages(suppressWarnings(
  readr::read_csv("~/Desktop/data/balances2-ed.csv", show_col_types = FALSE)
))
)

# Remove outliers
balances_clean <- dplyr::filter(
  balances,
  balance > mean(balance) - 3*sd(balance),
  balance < mean(balance) + 3*sd(balance)
)

# Compute summary stats (no range)
summary_stats <- balances_clean %>%
  summarise(
    `Total Preschools` = n(),
    `Mean Balance (R)` = mean(balance),
    `Median Balance (R)` = median(balance),
    `Minimum Balance (R)` = min(balance),
    `Maximum Balance (R)` = max(balance)
  ) %>%
  mutate(across(where(is.numeric), round, 0))

# Pretty table
summary_stats %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    table.border.top.width = 0,
    table.border.bottom.width = 0
  )

```

<br>

```{r balances-density, echo=FALSE}

#| warning: false
#| message: false

# Read in the data
balances <- read_csv("~/Desktop/data/balances2-ed.csv")

# Create histogram (one extreme outlier removed - user submitted the same income 3 times)
balances_clean <- dplyr::filter(
  balances,
  balance > mean(balance) - 3*sd(balance),
  balance < mean(balance) + 3*sd(balance)
)

ggplot(balances_clean, aes(x = balance)) +
  geom_histogram(binwidth = 3000, fill = "steelblue", colour = "white") +
  scale_x_continuous(labels = comma, breaks = pretty_breaks(n = 7)) +
  scale_y_continuous(labels = comma) +
  labs(
    x = "Balance (R)",
    y = "Number of Preschools"
  ) +
  theme_minimal()
```

<br>

##### Income and Expense Categories
_Percentage breakdown of income and expense categories during the selected period. Each pie shows the share of total income or total expenses contributed by each category._

::: {style="width:100%"}

``` {r expense-income-types, echo=FALSE}
#| warning: false
#| message: false

conflicted::conflicts_prefer(plotly::layout)

# --- Income ---
income_data <- read_csv("~/Desktop/data/income-types.csv") %>%
  arrange(desc(Percentage)) %>%
  mutate(Category = factor(Category, levels = Category))

p1 <- plot_ly(
  data = income_data,
  labels = ~Category,
  values = ~Percentage,
  type = "pie",
  sort = FALSE,
  direction = "clockwise",
  textinfo = "label+percent",
  textposition = "auto",
  hoverinfo = "text",
  text = ~paste(Category, ": ", Percentage, "%"),
  marker = list(
    colors = c("#1DBADF", "#FF2180", "#FFD525", "#83BB26", "#27385A"),
    line = list(color = "white", width = 2)
  ),
  domain = list(x = c(0, 0.48))  # left half
) %>%
  plotly::layout(
    showlegend = FALSE,
    uniformtext = list(minsize = 12, mode = "hide"),
    startangle = 90,
    margin = list(l = 0, r = 0, t = 80, b = 0),
    automargin = TRUE,
    autosize = TRUE,
    width = NULL,
    height = 500
  )

# --- Expenses ---
expense_data <- read_csv("~/Desktop/data/expense-types.csv") %>%
  arrange(desc(Percentage)) %>%
  mutate(Category = factor(Category, levels = Category))

p2 <- plot_ly(
  data = expense_data,
  labels = ~Category,
  values = ~Percentage,
  type = "pie",
  sort = FALSE,
  direction = "clockwise",
  textinfo = "label+percent",
  textposition = "auto",
  hoverinfo = "text",
  text = ~paste(Category, ": ", Percentage, "%"),
  marker = list(
    colors = c("#1DBADF", "#FF2180", "#FFD525", "#83BB26", "#27385A", "#8EDCEF", "#FF90BF"),
    line = list(color = "white", width = 2)
  ),
  domain = list(x = c(0.52, 1))  # right half
) %>%
  plotly::layout(
    showlegend = FALSE,
    uniformtext = list(minsize = 12, mode = "hide"),
    startangle = 90,
    margin = list(l = 0, r = 0, t = 80, b = 0),
    automargin = TRUE,
    autosize = TRUE,
    width = NULL,
    height = 500
  )

# --- Combine side by side ---
combined <- subplot(p1, p2, widths = c(0.5, 0.5)) %>%
  layout(
    margin = list(b = 120, t = 80),  # add bottom space
    annotations = list(
      list(
        text = "Income Types",
        x = 0.16,   # centre under left pie
        y = -0.10,  # move below chart
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 16)
      ),
      list(
        text = "Expense Types",
        x = 0.83,   # centre under right pie
        y = -0.10,
        xref = "paper",
        yref = "paper",
        showarrow = FALSE,
        font = list(size = 16)
      )
    )
  )

combined


```

:::

<br>

##### DBE Subsidy

##### Preschools Recording DBE Subsidies
_Among those who logged income and expenses._

::: {.stat-block}

::: {.stat-card}
<p class="stat-number">17</p>
<p class="stat-label">preschools logged DBE subsidy</p>
:::

::: {.stat-card}
<p class="stat-number">R 89,040</p>
<p class="stat-label">average DBE subsidy amount received</p>
:::

::: {.stat-card}
<p class="stat-number">R 45,793</p>
<p class="stat-label">median DBE subsidy amount</p>
:::

::: {.stat-card}
<p class="stat-number">50</p>
<p class="stat-label">children covered on average</p>
:::

::: {.stat-card}
<p class="stat-number">15</p>
<p class="stat-label">minimum number of children covered</p>
:::

::: {.stat-card}
<p class="stat-number">100</p>
<p class="stat-label">maximum number of children covered</p>
:::

:::

<br>

##### Preschool Fees
_Preschool fee information, based on the 49 preschools which have added income/expenses to the app._

::: {.stat-block}

::: {.stat-card}
<p class="stat-number">36</p>
<p class="stat-label">preschools logged preschool fees</p>
:::

::: {.stat-card}
<p class="stat-number">R 438</p>
<p class="stat-label">average preschool fee amount received per child</p>
:::

::: {.stat-card}
<p class="stat-number">70.1%</p>
<p class="stat-label">average % of children who paid preschool fees per month (out of all children at preschools where fees were logged)</p>
:::

:::

<br>

#### 4. Teaching & Learning

##### Lesson planning
The classroom activities and resources sections of the app have generated a lot of interest.

::: {.stat-block}

::: {.stat-card}
<p class="stat-number">221</p>
<p class="stat-label">users have used the lesson planning tool</p>
:::

::: {.stat-card}
<p class="stat-number">31</p>
<p class="stat-label">average number of days planned per practitioner (counting practitioners who have used the lesson planning tool at least once)</p>
:::

:::

##### Child Progress Tracking
_lorem ipsum_

::: {.stat-block}

::: {.stat-card}
<p class="stat-number">40</p>
<p class="stat-label">preschools have used the child progress tracking tool</p>
:::

:::

###### Child Developmental Outcomes by Subcategory

```{r progress2, echo=FALSE}

#| warning: false
#| message: false

# Read CSV
data <- read_csv("~/Desktop/data/child-outcomes-developmental-areas.csv")

# Convert to long format
data_long <- pivot_longer(data,
                          cols = c("Positive", "Negative", "Don't know"),
                          names_to = "Response",
                          values_to = "Count")

# ECD Connect colours
ecd_colours <- c("#1DBADF", "#FF2180", "#FFD525", "#83BB26", "#27385A", "#8EDCEF", "#FF90BF")

# Assign colours to responses
response_colours <- c("Positive" = "#83BB26",
                      "Negative" = "#FF2180",
                      "Don't know" = "#1DBADF")

# Horizontal stacked bar chart
ggplot(data_long, aes(y = SubcategoryName, x = Count, fill = Response)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = response_colours) +
  labs(title = "Child Developmental Outcomes by Subcategory",
       x = "Number of Skills",
       y = "Developmental Subcategory",
       fill = "Response") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 9),
        legend.position = "bottom")


```

---

Additional thoughts in Education. As we get a larger user base, particularly for the ECD Connect App, we are also able to send out comms via WA and learn more about our users, etc. -- so e.g. the points competition form & results.

---


### Health (CHW Connect)

#### 1. Client Demographics
_Demographics of all clients added since launch__

##### Child Gender Distribution

``` {r gg-children-gender, echo=FALSE }
#| warning: false
#| message: false

conflicted::conflicts_prefer(plotly::layout)

# Load data
gender_data <- read_csv("~/Desktop/data/gg-gender.csv")

# Prepare data
gender_data <- gender_data %>%
  arrange(desc(count)) %>%
  mutate(
    gender = factor(gender, levels = gender),
    Percentage = round(count / sum(count) * 100, 1)
  )

# Create interactive pie
plot_ly(
  data = gender_data,
  labels = ~gender,
  values = ~count,
  type = "pie",
  sort = FALSE,
  direction = "clockwise",
  textinfo = "label+percent",
  textposition = "auto",
  hoverinfo = "text",
  text = ~paste(gender, ": ", Percentage, "%"),
  marker = list(
    colors = c("#83BB26", "#1DBADF", "#FFD525"),
    line = list(color = "white", width = 2)
  )
) %>%
  layout(
  showlegend = FALSE,
  uniformtext = list(minsize = 12, mode = "hide"),
  startangle = 90,
  margin = list(l = 0, r = 0, t = 10, b = 0),
  automargin = TRUE,
  autosize = TRUE,
  height = NULL,
  width = NULL,
  piecolorway = c("#83BB26", "#1DBADF", "#FFD525"),
  annotations = list(),
  # 👇 this is what enables the leader (connector) lines
  showlegend = FALSE,
  # Important connector line setting:
  newshape = list(line = list(color = "grey", width = 1))
) %>%
  config(responsive = TRUE)


```

##### Birth Cohorts of Active Children

```{r birth-cohorts, echo=FALSE }
#| warning: false
#| message: false

# ---- Load your real data ----
df <- read_csv("~/Desktop/data/gg-ages.csv")

# Expecting columns:
# birth_period   count

# ---- ECD Connect colours ----
ecd_cols <- c(
  teal_blue = "#1DBADF",
  pink = "#FF2180",
  yellow = "#FFD525",
  green = "#83BB26",
  navy = "#27385A",
  light_teal = "#8EDCEF",
  light_pink = "#FF90BF"
)

# ---- Ordering: year desc, Jul–Dec above Jan–Jun ----
df <- df %>%
  mutate(
    year = as.integer(str_sub(birth_period, -4)),
    half = if_else(str_detect(birth_period, "Jul–Dec"), 2, 1),
    birth_period = fct_reorder(birth_period, year*10 + half)
  )

# ---- Plot ----
ggplot(df, aes(x = birth_period, y = count)) +
  geom_col(fill = ecd_cols["teal_blue"]) +
  geom_text(aes(label = count), vjust = -0.4, size = 3.2) +
  scale_y_continuous(expand = expansion(mult = c(0, .15))) +
  labs(
    x = "",
    y = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", colour = ecd_cols["navy"]),
    plot.subtitle = element_text(colour = ecd_cols["navy"]),
    axis.text.x = element_text(angle = 45, hjust = 1, colour = ecd_cols["navy"]),
    axis.title.y = element_text(colour = ecd_cols["navy"]),
    panel.grid.minor = element_blank()
  )
```


##### Number of Clients Added per Province
```{r province-clients, echo=FALSE}
#| warning: false
#| message: false

df <- read_csv("~/Desktop/data/clients-by-province.csv")

# Clean column names exactly as they appear in your CSV
df_clean <- df %>%
  rename(
    Province = Province,
    CHWs = CHWs,
    "Pregnant clients" = `Pregnant Clients Added`,
    "Child clients" = `Child Clients Added`
  ) %>%
  select(Province, CHWs, "Pregnant clients", "Child clients")

# Pivot to long format
df_long <- df_clean %>%
  pivot_longer(
    cols = c(CHWs, "Pregnant clients", "Child clients"),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(
    Metric = factor(Metric, levels = c("CHWs", "Pregnant clients", "Child clients")),
    Value = ifelse(Value == 0, 0.01, Value)  # optional – prevents invisible bars
  )

# ECD Connect colours
ecd_cols <- c(
  "CHWs" = "#1DBADF",     # teal blue
  "Pregnant clients" = "#FF2180", # bright pink
  "Child clients" = "#FFD525"     # yellow
)

ggplot(df_long, aes(x = Province, y = Value, fill = Metric)) +
  geom_col(position = position_dodge(width = 0.8)) +
  scale_fill_manual(values = ecd_cols) +
  labs(
    y = "Count",
    x = "Province",
    fill = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )

```


#### 2. Visits & Client Indicators
##### Visits Completed per Month (November 2024 to October 2025)

```{r gg-visits-bar, echo=FALSE}
#| message: false
#| warning: false
#| paged-print: true
#| out.width: 100%

visits <- read_csv("~/Desktop/data/visits.csv")

# Order months chronologically
visits$Month <- factor(visits$Month, levels = unique(visits$Month))

# Ensure Visit Type order (bottom first)
visits$`Visit Type` <- factor(
  visits$`Visit Type`,
  levels = c("Number of pregnant client visits", "Number of child visits")
)

# Stacked bar chart with reversed legend order
ggplot(visits, aes(x = Month, y = `Visits Completed`, fill = `Visit Type`)) +
  geom_bar(stat = "identity", colour = "white", width = 0.8) +
  scale_fill_manual(
    values = c(
      "Number of pregnant client visits" = "#FF2180",  # pink (top)
      "Number of child visits" = "#1DBADF"             # teal (bottom)
    ),
    breaks = c("Number of child visits", "Number of pregnant client visits") # ensures correct legend order
  ) +
  labs(
    x = NULL,
    y = "Visits Completed",
    fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.box.margin = margin(t = -10),
    legend.text = element_text(size = 10)
  )
```

<br>



##### Active CHWs and Average Visits per Month
_DRAFT narrative: An active CHW is one who completed at least one visit in that month. Pilots were conducted in 4 provinces over the past 2 years. A pilot in one province ended and CHWs are no longer using the app. A pilot in another province ended in November, with CHWs being deactivated._

```{r gg-visits-chws, echo=FALSE}
#| message: false
#| warning: false
#| paged-print: true

data <- read_csv("~/Desktop/data/avg-visits-chws.csv")

# Convert Month to date format if it's not already
data <- data %>%
  mutate(Month = as.Date(Month))

# Prefer plotly's layout
conflicted::conflicts_prefer(plotly::layout)

# Create interactive line chart
plot_ly(data) %>%
  add_lines(
    x = ~Month,
    y = ~`Number of active CHWs`,
    name = "Active CHWs",
    line = list(color = "#1DBADF", width = 3)
  ) %>%
  add_lines(
    x = ~Month,
    y = ~`Avg number of visits per active CHW`,
    name = "Avg Visits per CHW",
    line = list(color = "#FF2180", width = 3, dash = "dot")
  ) %>%
  layout(
    xaxis = list(title = "Month"),
    yaxis = list(title = "Count / Average"),
    legend = list(x = 0.05, y = 0.95)
  )

```


##### Stunting Prevalance by Month
_The Grow Great team is able to view stunting rates alongside the number of child visits completed; as well as filtered by province/district, etc. The graph below is showing stunting prevalence by month, across all areas._

```{r stunting, echo=FALSE}

#| warning: false
#| message: false

# Read CSV
stunting <- read_csv("~/Desktop/data/stunting.csv")

# Plot

ggplot(stunting, aes(x = Month, y = `Stunting Prevalence %`)) +
  geom_col(fill = "#1DBADF") +  # Teal blue
  geom_text(aes(label = paste0(`Stunting Prevalence %`, "%")),
            vjust = -0.3, size = 3.5, colour = "#27385A") +  # Navy text
  labs(
    x = NULL,
    y = "Stunting Prevalence (%)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(colour = "#27385A", face = "bold"),
    axis.text.x = element_text(colour = "#27385A"),
    axis.text.y = element_text(colour = "#27385A"),
    axis.title.y = element_text(colour = "#27385A"),
    panel.grid.minor = element_blank()
  )
```

##### Potential Developmental Concerns Identified
_Percentage of screenings where milestone was not achieved_

``` {r child-development, echo=FALSE}
#| warning: false
#| message: false

df <- read_csv("~/Desktop/data/developmental-screening.csv")

# compute follow-up rate
rates <- df %>%
  pivot_wider(names_from = Response, values_from = Count) %>%
  mutate(
    total = `On track` + `Needs follow-up`,
    followup_rate = `Needs follow-up` / total
  )

ggplot(rates, aes(x = `Developmental Area`, y = followup_rate)) +
  geom_col(fill = "#FF2180") +   # ECD Connect bright pink
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = NULL,
    y = "Needs follow-up (%)"
  ) +
  theme_minimal(base_size = 15)
```

##### Referrals (November 2024 to November 2025)
_Referrals are automatically raised by the app during visits. After the visit, the CHW checks the referral box to indicate whether the referral to the clinic has been made_

::: {.stat-block}

::: {.stat-card}
<p class="stat-number">16,350</p>
<p class="stat-label">referrals raised</p>
:::

::: {.stat-card}
<p class="stat-number">6,681</p>
<p class="stat-label">referrals raised</p>
:::

::: {.stat-card}
<p class="stat-number">2,338</p>
<p class="stat-label">back-referrals completed (feedback from the clinic)</p>
:::

:::



``` {r referrals, echo=FALSE}
#| warning: false
#| message: false


child_df_raw <- read_csv("~/Desktop/data/child-referrals.csv") %>%
  rename(
    Referral.type = `Referral type`,
    Total = Total
  )

# compute the max *before* mutate
max_total <- max(child_df_raw$Total, na.rm = TRUE)

child_df <- child_df_raw %>%
  mutate(
    Bar = purrr::map_chr(
      Total,
      ~ paste(rep("▉", round(.x / max_total * 10)), collapse = "")
    )
  )

gt(child_df) %>%
  gt::text_transform(
    locations = cells_body(columns = Bar),
    fn = function(x) paste0("<span style='color:#1DBADF'>", x, "</span>")
  ) %>%
  tab_header(
    title = "Top 5 Referral Reasons – Children",
    subtitle = "Nov 2024 – Nov 2025"
  ) %>%
  cols_label(
    Referral.type = "Referral reason",
    Total = "Count",
    Bar = "Visual scale"
  ) %>%
tab_options(table.width = pct(100)) %>%
cols_width(Bar ~ px(120)) %>%
tab_style(
  style = cell_text(whitespace = "nowrap"),
  locations = cells_column_labels(columns = Bar)
)

```

``` {r referrals-pregnant, echo=FALSE}
#| warning: false
#| message: false

child_df <- read_csv("~/Desktop/data/pregnant-referrals.csv") %>%
  rename(Referral.type = `Referral type`, Total = Total) %>%
  mutate(
    Bar = purrr::map_chr(
      Total,
      ~ paste(rep("▉", round(.x / max(child_df$Total) * 20)), collapse = "")
    )
  )

gt(child_df) %>%
  gt::text_transform(
    locations = cells_body(columns = Bar),
    fn = function(x) paste0("<span style='color:#1DBADF'>", x, "</span>")
  ) %>%
  tab_header(
    title = "Top 5 Referral Reasons – Pregnant Clients",
    subtitle = "Nov 2024 – Nov 2025"
  ) %>%
  cols_label(
    Referral.type = "Referral reason",
    Total = "Count",
    Bar = "Visual scale"
  ) %>%
tab_options(table.width = pct(100)) %>%
cols_width(Bar ~ px(120)) %>%
tab_style(
  style = cell_text(whitespace = "nowrap"),
  locations = cells_column_labels(columns = Bar)
)

```



