---
title: "ECD Connect Data Report"
author: "Kim Tichmann"
format: html
---

# October 2025 Report

## 1. Reach and Scale of ECD Connect
A summary of ECD Connect's reach.

This includes the ECD Connect App (open-access), all ECD Connect Partner versions (white-label) and Grow Great's CHW Connect.

::: {.stat-block}

::: {.stat-card}
<p class="stat-number">4</p>
<p class="stat-label">partner organisations</p>
:::

::: {.stat-card}
<p class="stat-number">1,912</p>
<p class="stat-label">unique user sign-ups</p>
:::

::: {.stat-card}
<p class="stat-number">12,300</p>
<p class="stat-label">total number of children added across all platforms</p>
:::

:::

```{r beneficiaries-map, echo=FALSE}
#| message: false
#| warning: false
#| paged-print: true

library(sf)
library(dplyr)
library(ggplot2)
library(conflicted)
library(stringr)
library(patchwork)
library(scales)

conflict_prefer("filter","dplyr")

load("sub_places.rda")  
load("postalcodes_geo.rda")

df_children <- read.csv("codes-ed-13-10-2025.csv", stringsAsFactors = FALSE)
df_clients  <- read.csv("GG-location-13-10-2025.csv", stringsAsFactors = FALSE)

districts <- suppressMessages(suppressWarnings(st_read("DistrictMunicipalities2018_Final.shp", quiet = TRUE)))

df_clients2 <- df_clients %>%
  mutate(
    District_clean = trimws(toupper(District))
  )

districts2 <- districts %>%
  mutate(
    DISTRICT_NAME_clean = trimws(toupper(DISTRICT_N))
  )

# Which CSV district names are in the map?
matched <- df_clients2 %>%
  filter(District_clean %in% districts2$DISTRICT_NAME_clean)
invisible(nrow(matched))

# ensure the district polygon column exists
if (!"DISTRICT_NAME_clean" %in% names(districts2)) {
  stop("Expected DISTRICT_NAME_clean in 'districts2' but not found.")
}

# Join the health data to the district polygons
districts_health <- districts2 %>%
  left_join(df_clients2, by = c("DISTRICT_NAME_clean" = "District_clean"))

df_children2 <- df_children %>%
  mutate(PostalCode = as.character(PostalCode),
         TotalChildren = as.numeric(TotalChildren))

postalcodes_geo2 <- postalcodes_geo %>%
  mutate(PostalCode = as.character(code))

# join lat/lon + totals
pcs <- postalcodes_geo2 %>%
  left_join(df_children2, by = "PostalCode") %>%
  filter(!is.na(TotalChildren))

# convert to spatial points
pcs_sf <- st_as_sf(pcs, coords = c("lon", "lat"), crs = 4326, remove = FALSE)

# align CRS + spatial join to districts
pcs_sf <- st_transform(pcs_sf, st_crs(districts2))
pcs_with_district <- st_join(pcs_sf, districts2 %>% select(DISTRICT_NAME_clean), join = st_within)

# aggregate postal codes up to district
districts_ed <- pcs_with_district %>%
  st_set_geometry(NULL) %>%
  group_by(DISTRICT_NAME_clean) %>%
  summarise(TotalChildren = sum(TotalChildren, na.rm = TRUE)) %>%
  right_join(districts2, by = "DISTRICT_NAME_clean") %>%
  st_as_sf()

# ---- COMBINE BOTH LAYERS ----
health_for_map <- districts_health %>%
  select(DISTRICT_NAME_clean, geometry, Value = TotalClients) %>%
  mutate(Source = "Health")

ed_for_map <- districts_ed %>%
  select(DISTRICT_NAME_clean, geometry, Value = TotalChildren) %>%
  mutate(Source = "Education")

combined_map <- bind_rows(health_for_map, ed_for_map) %>%
  group_by(DISTRICT_NAME_clean) %>%
  summarise(
    TotalBeneficiaries = sum(Value, na.rm = TRUE),
    geometry = sf::st_union(geometry)
  ) %>%
  ungroup()

# --- Merge with district polygons to ensure full geometry coverage ---
combined_map <- districts2 %>%
  left_join(
    st_drop_geometry(combined_map),
    by = "DISTRICT_NAME_clean"
  )

# --- Plot ---
library(ggplot2)
library(viridis)

combined_map <- combined_map %>%
  mutate(
    BeneficiaryBand = cut(
      TotalBeneficiaries,
      breaks = c(-Inf, 0, 2000, 4000, 6000, Inf),
      labels = c("0", "1–2000", "2001–4000", "4001–6000", "6000+")
    )
  ) %>%
  mutate(
    BeneficiaryBand = factor(
      BeneficiaryBand,
      levels = c("0", "1–2000", "2001–4000", "4001–6000", "6000+")
    )
  )

ggplot(combined_map) +
    geom_sf(aes(fill = BeneficiaryBand), colour = "#fff", size = 0.15) +
  scale_fill_manual(
    values = c(
      "0" = "#C9CFD2",
      "1–2000" = "#8EDCEF",
      "2001–4000" = "#1DBADF",
      "4001-6000" = "#52607B",
      "6000+" = "#27385A"
    ),
    name = "Total beneficiaries",
    drop = FALSE
  ) +
  labs(
    title = "Total Beneficiaries by District",
    subtitle = "Combined Health and Education data",
    caption = "‘Beneficiaries’ refers to children (education and health) \n and pregnant clients (health) registered in the app."
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

```

## 2. Engagement and Activity
Users are engaging with the ECD Connect tools and submitting data.

### Education
#### ECD Connect App
_Since launch on 1 April 2025 to 13 October 2025_

::: {.stat-block}

::: {.stat-card .app-background}
<p class="stat-number">1,377</p>
<p class="stat-label">unique sign-ups</p>
:::

::: {.stat-card .app-background}
<p class="stat-number">464</p>
<p class="stat-label">preschools added</p>
:::

::: {.stat-card .app-background}
<p class="stat-number">1,434</p>
<p class="stat-label">children registered</p>
:::

:::

```{r app-attendance-bar, echo=FALSE}
#| message: false
#| warning: false
#| paged-print: true
#| out.width: 100%

df_appattendance <- read.csv("app-attendance.csv", stringsAsFactors = FALSE)

df_appattendance <- df_appattendance %>%
    mutate(
      Month = as.Date(paste0(Month,"-01")),
      Month = format(Month, "%B %Y")
    )

ggplot(data = df_appattendance, aes(x = Month, y = Registers.Saved)) +
  geom_bar(stat = "identity",  fill = "#FF2180")

```

#### ECD Connect Partner

### Health (CHW Connect)


